Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-04-09 SupportingInformation.Rmd
answering Reid's comment and adding notes for DAN
code edited for cedar creek : include 168 plots as per Tilman's paper, species name screened

---
title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

@@

<!--reading ceder creek data and generating raw and copula plots with it-->
```{r read_ceder_creek_data, echo=F, results="hide"}
set.seed(seed)
#-------------------------------------------------
# This function takes 
# Input :
#       db : a dataframe : smaller subset of ple_120 data to calculate avg. biomass for each plot
#       dp : a dataframe : smaller subset of pce_120 data to calculate avg. shannon's H for those plots
#       year : the year for which you want to calculate biomass and shannon's index data
#       rawplot : logical : T for rawdata plot, F for copula plot
#       ploton : logical for optional plotting

@@

cop_HB<-function(db,dp,year,rawplot,ploton){
  # db : biomass dataset for a specific year
  # dp : percent cover dataset for a specific year

  plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
  db<-subset(db,Plot%in%plots)
  dp<-subset(dp,Plot%in%plots)

@@

    # calculate shannon's index H for each plot (averaged over all Quadrates for that plot)
    x<-subset(dp,Plot==plots[i])
    uniq_sp<-sort(as.character(unique(x$Species)))  # These are the species in a particular plot
    Q<-as.character(unique(x$Quadrat))

    tab_each_sp<-data.frame(sp=uniq_sp,avg_cov=0,rel_avg_cov=0)

    for(sp in 1:length(uniq_sp)){
      my_sp<-uniq_sp[sp]
      tab_temp<-data.frame(Q=Q,abs_cov=0)
      ind<-which(x$Species==my_sp)
      li<-length(ind)
      tab_temp$abs_cov[1:li]<-x$Abscover[ind]
      tab_each_sp$avg_cov[[sp]]<-mean(tab_temp$abs_cov) # avg. over 4Quadrates for a sp.
    }

    tab_each_sp$rel_avg_cov<-tab_each_sp$avg_cov/sum(tab_each_sp$avg_cov)
    tab_dbp$avg.H[i]<- -sum((tab_each_sp$rel_avg_cov)*log(tab_each_sp$rel_avg_cov))

  tab_dbp<-na.omit(tab_dbp)

@@

a<-as.data.frame(table(sort(r_pce$Species)))
b<-as.character(a$Var1)
bad_sp<-c(b[51:53],b[156:158])

ind<-which(r_pce$Species%in%bad_sp)
r_pce_clean<-r_pce[-ind,]
#--------------------------------------------------------------------------------------------
pdf("./Results/Ceder_creek_results/raw_HB_all_yr.pdf",height = 6,width = 10)
op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
for(i in 1:length(yr)){
  db<-subset(r_ple,Year==yr[i])
  dp<-subset(r_pce_clean,Year==yr[i])

  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
}
par(op)
dev.off()
#-----------------------
#raw data plot for year2000
pdf("./Results/Ceder_creek_results/raw_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
saveRDS(sr,"./Results/Ceder_creek_results/raw_cop_cc_2000.RDS")
dev.off()
#----------------------------------------------------------------------------------------
cop_HB_all_yr<-vector("list",length(yr))
names(cop_HB_all_yr)<-yr

@@

  s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
  cop_HB_all_yr[[i]]<-s$cop_tab_dbp[,2:3]
}
par(op)
dev.off()
#-------------------------
#cop data plot for year2000
pdf("./Results/Ceder_creek_results/cop_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
cc_cop_2000<-s$cop_tab_dbp[,2:3]
dev.off()
#------------------

>>>>

---
title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, Daniel C. Reuman"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

@@

<!--reading ceder creek data and generating raw and copula plots with it-->
```{r read_ceder_creek_data, echo=F, results="hide"}
set.seed(seed)
#-------------------------------------------------
# This function takes 
# Input :
#       db : a dataframe : smaller subset of ple_120 data to calculate avg. biomass for each plot
#       dp : a dataframe : smaller subset of pce_120 data to calculate avg. shannon's H for those plots
#       plots : a vector of given 168 plot ids as used in reference papers by Tilman
#       year : the year for which you want to calculate biomass and shannon's index data
#       rawplot : logical : T for rawdata plot, F for copula plot
#       ploton : logical for optional plotting

@@

cop_HB<-function(db,dp,plots,year,rawplot,ploton){
  # db : biomass dataset for a specific year
  # dp : percent cover dataset for a specific year

  db<-subset(db,Plot%in%plots)
  dp<-subset(dp,Plot%in%plots)

@@

   # calculate shannon's index H for each plot (averaged over all Quadrates for that plot)
   x<-subset(dp,Plot==plots[i])
   x$Species<-trimws(x$Species) # to remove duplicated names of sp. (cutting white spaces) : e.g. 1999 : plot no 339
   uniq_sp<-sort(as.character(unique(x$Species)))  # These are the species in a particular plot
   Q<-as.character(unique(x$Quadrat))

   tab_each_sp<-data.frame(sp=uniq_sp,avg_cov=0,rel_avg_cov=0)
   #cat("=====================================plot : ",plots[i],"===========================\n")
   for(sp in 1:length(uniq_sp)){
    my_sp<-uniq_sp[sp]
    #cat("---------- sp = ",my_sp,"--------------------\n")
    tab_temp<-data.frame(Q=Q,abs_cov=0)
    ind<-which(x$Species==my_sp)
    li<-length(ind)
    #print(li)
    tab_temp$abs_cov[1:li]<-x$Abscover[ind]
    tab_each_sp$avg_cov[[sp]]<-mean(tab_temp$abs_cov) # avg. over 4Quadrates for a sp.
   }

  tab_each_sp$rel_avg_cov<-tab_each_sp$avg_cov/sum(tab_each_sp$avg_cov)
  tab_dbp$avg.H[i]<- -sum((tab_each_sp$rel_avg_cov)*log(tab_each_sp$rel_avg_cov))
 }
  tab_dbp<-na.omit(tab_dbp)

@@

a<-as.data.frame(table(sort(r_pce$Species)))
b<-as.character(a$Var1)
bad_sp<-c(b[51:53],b[156:158],"Disturbance") 
# bad sp : "Bare ground","Bare ground","BARE GROUND","Miscellaneous litter","Miscellaneous litter","MISCELLANEOUS LITTER"            
ind<-which(r_pce$Species%in%bad_sp)
r_pce_clean<-r_pce[-ind,]

#----------------------------------------------------------------------------------------------
# This sections selects 168 plots considered for both biomass and percent cover datasets in Tilman's papers 2001, 2006
i<-5#for Yr=2000
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
#------------------------------------------------------
id_plot_nosp<-which(db$NumSp==0 & db$Biomass==0) # These plots have no sp. seeded initially and have no biomass data
plots_no_biomass<-sort(unique(db$Plot[id_plot_nosp]))
plots_no_biomass

plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
length(plots)

excluded_plots<-intersect(plots,plots_no_biomass)

plots_168<-setdiff(plots,excluded_plots)

#--------------------------------------------------------------------------------------------
pdf("./Results/Ceder_creek_results/raw_HB_all_yr.pdf",height = 6,width = 10)
op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
for(i in 1:length(yr)){
  db<-subset(r_ple,Year==yr[i])
  dp<-subset(r_pce_clean,Year==yr[i])

  sr<-cop_HB(db=db,dp=dp,year=yr[i],plots=plots_168,rawplot=T,ploton = T)
}
par(op)
dev.off()
#-----------------------
#raw data plot for year2000
pdf("./Results/Ceder_creek_results/raw_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
sr<-cop_HB(db=db,dp=dp,year=yr[i],plots=plots_168,rawplot=T,ploton = T)
saveRDS(sr,"./Results/Ceder_creek_results/raw_cop_cc_2000.RDS")
dev.off()
#----------------------------------------------------------------------------------------
cop_HB_all_yr<-vector("list",length(yr))
names(cop_HB_all_yr)<-yr

@@

  s<-cop_HB(db=db,dp=dp,year=yr[i],plots=plots_168,rawplot=F,ploton = T)
  cop_HB_all_yr[[i]]<-s$cop_tab_dbp[,2:3]
}
par(op)
dev.off()
#-------------------------
#cop data plot for year2000
pdf("./Results/Ceder_creek_results/cop_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
s<-cop_HB(db=db,dp=dp,year=yr[i],plots=plots_168,rawplot=F,ploton = T)
cc_cop_2000<-s$cop_tab_dbp[,2:3]
dev.off()
#------------------
